Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-12-30 Paper/ms.tex
A bit more prior editing

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
prior knowledge, rather than the judgement of an informed expert on extrasolar
planets. Uniform distributions were used for parameters such as phases, where
time-translation symmetry seems plausible.
Truncated Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe relative to other possible ``informative'' assignments such as normal distributions.
For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
as $\approx e^{21}$ years, a very generous range. Truncation intervals are
specified with the $T(a,b)$ notation in Table~\ref{tab:priors}.

An apparently strange choice is the conditional prior for the (logarithms of)
the orbital
periods, which is a biexponential distribution given a location parameter
$\mu_P$ and scale parameter, or width, $w_P$.
Rather than assigning independent priors to the log periods, the hierarchical
model allows for the periods to ``cluster around'' a typical period $\mu_P$ if
there is evidence for this. On the other hand, independent priors for the
periods would imply a strong prior commitment to the hypothesis that the
periods are spread out across the whole prior volume (equivalent to assuming
a fixed large value for $w_P$.
A more conventional choice for the conditional prior given $\mu_P$ and $w_P$
would have been a normal distribution. However, the \citet{rjobject} software
needs to know the corresponding cumulative distribution and its inverse, which
are not available in closed form for the normal distribution.

@@

\begin{table*}
\begin{tabular}{|l|l|l|}
\hline
Quantity	&	Meaning		& Prior\\
\hline
{\bf Hyperparameters}	&	\\
$N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
$\mu_P$		&	Median orbital period	& $\log(\mu_P) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
$w_P$		&	Diversity of orbital periods & $w_P \sim$ Uniform$(0.1, 3)$\\
$\mu_A$		&	Mean amplitude	& $\log(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
\hline
{\bf Planet Parameters}\\
$P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential($\log(\mu_P), w_P$)\\
$A_i$		&	Semi-mplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
$\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
$e_i$		&	Eccentricity of orbit	&	Beta$(1, 3.1)T(0, 0.8)$\\
$\omega_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
\hline
{\bf Other}\\
$m_0$		&	Constant DC offset	&	Uniform$(m_{0, \rm min}, m_{0, \rm max})$\\
$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
$\nu$		& Shape parameter for $t$-distribution for noise & $\log(\nu) \sim$ Uniform$(\log(0.01), \log(1000))$\\
\hline
{\bf Data}\\
$Y_i$		& Radial velocity measurements	&
		Student-$t\left(m(t_i), \sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}, \nu\right)$
\end{tabular}
\caption{All of the prior distributions in our Bayesian model.
The priors for the planet parameters are defined conditional on the values
of the hyperparameters. Uniform priors were used for parameters like phases.
For parameters where a rough initial guess is possible, heavy-tailed Cauchy
distributions were used so this information could be taken into account
in a non-dogmatic way. Time units are in days and amplitude units are
in metres per second. The maximum number of planets, $N_{\rm max}$, was set
to 10. The prior limits for $m_0$ were set to the minimum and maximum value
in the dataset, which is not strictly a valid strategy.\label{tab:priors}}
\end{table*}

>>>>

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
prior knowledge, rather than the judgement of an informed expert on extrasolar
planets. Uniform distributions were used for parameters such as phases, where
time-translation symmetry seems plausible. For some parameters we assigned the
distribution in terms of the log of the parameter, rather than the parameter
itself, when the parameter is positive and uncertain by orders of magnitude.
Truncated Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe relative to other possible ``informative'' assignments such as normal distributions.
For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year but could be as low as $\approx e^{-21}$ years or as high
as $\approx e^{21}$ years, a very generous range. A uniform distribution for
$\log(\mu_P)$ would have been more conventional, whereas the Cauchy distribution
expresses a slight preference for $\mu_P$ being of order one year.

An apparently strange choice is the conditional prior for the (logarithms of)
the orbital
periods, which is a biexponential distribution given a location parameter
$\mu_P$ and scale parameter, or width, $w_P$.
Rather than assigning independent priors to the log periods, the hierarchical
model allows for the periods to ``cluster around'' a typical period $\mu_P$ if
there is evidence for this. On the other hand, independent priors for the
periods would imply a strong prior commitment to the hypothesis that the
periods are spread out across the whole prior volume (equivalent to assuming
a fixed large value for $w_P$).
A more conventional choice for the conditional prior given $\mu_P$ and $w_P$
would have been a normal distribution. However, the \citet{rjobject} software
needs to know the corresponding cumulative distribution and its inverse, which
are not available in closed form for the normal distribution.

@@

\begin{table*}
\begin{tabular}{|l|l|l|}
\hline
Quantity	&	Meaning		& Prior\\
\hline
{\bf Hyperparameters}	&	\\
$N$		& Number of planets	& Uniform$(\{0, 1, ..., N_{\rm max}\})$\\
$\mu_P$		&	Median orbital period	& $\log(\mu_P) \sim$ Cauchy$(5.9, 1)T(-15.3, 27.1)$\\
$w_P$		&	Diversity of orbital periods & $w_P \sim$ Uniform$(0.1, 3)$\\
$\mu_A$		&	Mean amplitude	& $\log(\mu_A) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
\hline
{\bf Planet Parameters}\\
$P_i$		&	Orbital period	&	$\log(P_i) \sim $ Biexponential($\log(\mu_P), w_P$)\\
$A_i$		&	Semi-mplitude of signal	& Exponential$(\textnormal{mean}=\mu_A)$\\
$\phi_i$	&	Phase of signal	&	Uniform($0, 2\pi$)\\
$e_i$		&	Eccentricity of orbit	&	Beta$(1, 3.1)T(0, 0.8)$\\
$\omega_i$	&	Viewing angle	&	Uniform$(0, 2\pi)$\\
\hline
{\bf Other}\\
$m_0$		&	Constant DC offset	&	Uniform$(m_{0, \rm min}, m_{0, \rm max})$\\
$\sigma_{\rm extra}$	& ``Extra noise'' parameter	& $\log(\sigma) \sim$ Cauchy$(0, 1)T(-21.2, 21.2)$\\
$\nu$		& Shape parameter for $t$-distribution for noise & $\log(\nu) \sim$ Uniform$(\log(0.01), \log(1000))$\\
\hline
{\bf Data}\\
$Y_i$		& Radial velocity measurements	&
		Student-$t\left(m(t_i), \sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}, \nu\right)$
\end{tabular}
\caption{All of the prior distributions in our Bayesian model.
The priors for the planet parameters are defined conditional on the values
of the hyperparameters. Uniform priors were used for parameters like phases.
For parameters where a rough initial guess is possible, heavy-tailed Cauchy
distributions were used so this information could be taken into account
in a non-dogmatic way. Time units are in days and amplitude units are
in metres per second. The maximum number of planets, $N_{\rm max}$, was set
to 10. The prior limits for $m_0$ were set to the minimum and maximum value
in the dataset, which is not strictly a valid strategy. The 
Truncation intervals are specified with the $T(a,b)$ notation.\label{tab:priors}}
\end{table*}
